Hierarchy Builder

### 2020-04-27 coqws/easychair.tex
update coqws paper raping TeX

\section{Context and Talk proposal}

@@

\section{A hierarchy with a diamond in 50 lines}

\begin{coqcode}
HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid *)
  zero  : A;
  add   : A -> A -> A;
  addrA : associative add;
  addrC : commutative add;
  add0r : left_id zero add;
}.

HB.structure Definition CMonoid := { A of CMonoid_of_Type A }. (* The structure thereof *)

Notation "0" := zero.
Infix    "+" := add.

Check 0.     (* ?M : CMonoid.type |- 0 : ?M *)
Check addrC. (* ?M : CMonoid.type |- forall x y : ?M, x + y = y + x *)
\end{coqcode}

@@

\begin{wrapfigure}[0]{r}{.4\textwidth}
  \includegraphics[width=.4\textwidth]{diagram.pdf}
  %\caption{The hierarchy}
  %\label{fig:hierarchy}
\end{wrapfigure}
\
\begin{coqcode}
HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
  opp   : A -> A;
  addNr : left_inverse zero opp add; (* add is available here *)
  addrN : right_inverse zero opp add;
}.

HB.structure Definition AbelianGrp :=
  { A of AbelianGrp_of_CMonoid A & }.

Notation "- x" := (opp x).

Check forall G : AbelianGrp.type, forall x : G, x + - x = 0.

HB.mixin Record SemiRing_of_CMonoid A of CMonoid A := {
  one    : A;
  mul    : A -> A -> A;
  mulrA  : associative mul;
  mul1r  : left_id one mul;
  mulr1  : right_id one mul;
  mulrDl : left_distributive mul add;
  mulrDr : right_distributive mul add;
  mul0r  : left_zero zero mul;
  mulr0  : right_zero zero mul;
}.

HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & }.

Notation "1" := one.
Infix    "*" := mul.

Check forall S : SemiRing.type, forall x : S, x * x = 0 + 1. (* can I use - here? *)

HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.

Check forall R : Ring.type, forall x : R, x * - (1 + x) = 0 + 1.

Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
HB.instance Z Z_CMonoid.
Definition Z_AbelianGrp := AbelianGrp_of_CMonoid.Build Z Z.opp Z.add_opp_diag_l.
HB.instance Z Z_AbelianGrp.
Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l ...
HB.instance Z Z_SemiRing.

Check forall x : Z, x * - (1 + x) = 0 + 1.
\end{coqcode}

For space constraints we omit the declaration of the green arrow
in the picture (about 30 lines) which is an alternative interface
to build examples of the \coq{Ring} structure by providing all
proofs at once. In that case the user is spared to prove the commutativity
of the Monoid operation since it is implied by the other axioms~\cite{nearrings}.

\bibliography{biblio}

>>>>

\paragraph{Context and Talk proposal}

@@

\paragraph{A hierarchy with a diamond in 50 lines}

\begin{coqcode}
HB.mixin Record CMonoid_of_Type A := { (* The set of axioms making A a commutative monoid. *)
  zero  : A;
  add   : A -> A -> A;
  addrA : associative add;  (* `add` is associative  *)
  addrC : commutative add;  (* `add` is commutative  *)
  add0r : left_id zero add; (* `zero` is the neutral *)
}.
HB.structure Definition CMonoid := { A of CMonoid_of_Type A }. (* The structure thereof. *)
Notation "0" := zero.
Infix    "+" := add.

(* The type of the operations and axioms depend on a CMonoid.type structure. *)
Check addrC. (* ?M : CMonoid.type |- forall x y : ?M, x + y = y + x *)

HB.mixin Record AbelianGrp_of_CMonoid A of CMonoid A := {
  (* We can write `add` here since A is a  CMonoid   *)
  opp   : A -> A;
  addNr : left_inverse zero opp add; (* `opp` is the additive inverse *)
}.
\end{coqcode}

@@

\begin{wrapfigure}[0]{r}{.35\textwidth}
  \vspace{1em}
  \includegraphics[width=.35\textwidth]{diagram.pdf}
  \vspace{-1.5em}
  %\caption{
    \center{\small{Figure 1: The hierarchy}}
   % }
  \label{fig:hierarchy}
\end{wrapfigure}
\
\begin{coqcode}

HB.structure Definition AbelianGrp := { A of AbelianGrp_of_CMonoid A & }.
Notation "- x" := (opp x).
Notation "x - y" := (add x (opp y)).

(* A statement in the signature of an Abelian group G. *)
Check forall G : AbelianGrp.type, forall x : G, x - x = 0.

HB.mixin Record SemiRing_of_CMonoid A of CMonoid A := {
  one    : A;
  mul    : A -> A -> A;
  mulrA  : associative mul;  (* `mul` is associative   *)
  mul1r  : left_id one mul;  (* `one` is left neutral  *)
  mulr1  : right_id one mul; (* `one` is right neutral *)
  mulrDl : left_distributive mul add;  (* `mul` distributes over *)
  mulrDr : right_distributive mul add; (*   `add` on both sides  *)
  mul0r  : left_zero zero mul;  (* `zero` is absorbing `mul`     *)
  mulr0  : right_zero zero mul; (*   on both sides               *)
}.
HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & }.
Notation "1" := one.
Infix    "*" := mul.

(* A statement in the signature of a Semiring S.  *)
Check forall S : SemiRing.type, forall x : S, x * 1 + 0 = x.

(* We need no additional mixin to declare the structure of rings. *)
HB.structure Definition SemiRing := { A of SemiRing_of_CMonoid A & AbelianGrp_of_CMonoid A & }.

(* A statement in the signature of a Ring R.  *)
Check forall R : Ring.type, forall x : R, x * -1 = - x.

(* We equip Z with the structures of CMonoid, AbelianGrp and SemiRing; HB equips it with Ring. *)
Definition Z_CMonoid    := CMonoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_comm Z.add_0_l.
HB.instance Z Z_CMonoid.
Definition Z_AbelianGrp := AbelianGrp_of_CMonoid.Build Z Z.opp Z.add_opp_diag_l.
HB.instance Z Z_AbelianGrp.
Definition Z_SemiRing   := SemiRing_of_CMonoid.Build Z 1%Z Z.mul Z.mul_assoc Z.mul_1_l ...
HB.instance Z Z_SemiRing.

(* A statement in the signature of the Z ring *)
Check forall x : Z, x * - (1 + x) = 0 + 1.
\end{coqcode}

For space constraints we omit the declaration
of the green arrows (Figure 1) which are alternative interfaces
called factories. They provide specialized interfaces to build structures for others.
Factories are used to accommodate the insertion of
intermediate structures in a hierarchy without breaking user code, as we all
as to implement shortcuts: the factory to build a \coq{Ring} from \coq{Type}
(rightmost green arrow)
can omit the one for the commutativity of the addition, since it is implied
by the other axioms~\cite{nearrings}.
\bibliography{biblio}
